---
title: "Papp Preliminary Analysis"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---

Load packages.

```{r}
library(readxl)
library(tidyverse)
library(knitr)
library(ggplot2)
library(cowplot)
```

Read in ground cover data.

```{r}
papp_ground <- read_xlsx(path = "PappFieldData.xlsx", sheet = 2)
papp_species <- read.csv("Species_List.csv")
papp_ground <- left_join(papp_ground, papp_species, by = "Species")

# Change order of ecological conditions
papp_ground$EcoCondition <- factor(papp_ground$EcoCondition, 
                                   levels = c("Disturbed", "Undisturbed", "Ecologically Significant"))

```

# Canopy Density Analysis

* Collapse the data set so we have one canopy density value per plot

```{r}
canopy_dens <- 
  papp_ground %>%
  group_by(Site, Plot_Num, Location, EcoCondition) %>%
  summarise(Canopy_Density = unique(Canopy_Density),
            Canopy_Density_Adj_Trail = unique(Canopy_Density_Adjacent_Trail))
```

## Visualizing the data

* Plot Canopy Density using facets to split by site and color to separate by ecocondition

```{r}
ggplot(data = canopy_dens, aes(x = Location, y = Canopy_Density, fill = EcoCondition)) +
  geom_boxplot() +
  facet_grid(Site~.) +
  ylab("Canopy Density") + 
  scale_fill_discrete(name = "Ecological\nCondition",
                      breaks = c("Disturbed", "Undisturbed", "Ecologically Significant"),
                      labels = c("Disturbed", "Undisturbed", "Ecologically\nSignificant")) +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-3-CanDens-Loc-Site.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

* Plot Canopy Density of Adjacent Trails using ggplot to split by site

```{r}
ggplot(data = canopy_dens, aes(x = Site, y = Canopy_Density_Adj_Trail)) +
  geom_boxplot() +
  ylab("Canopy Density on Trail") + 
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-4-CanDens-Adj.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

## Statistical Analysis

* Perform an ANOVA to check for differences in Canopy Density at Sites, Locations, and EcoConditions

```{r}
canopy_dens_lm_all <- lm(data = canopy_dens, Canopy_Density ~ Site * Location * EcoCondition )
kable(anova(canopy_dens_lm_all), digits = 3, format = "pandoc", caption = "ANOVA table of Canopy Density")
```

* Perform a t-test to check for differences in Canopy Density at Adjacent Trails

```{r}
with(canopy_dens, kable(t.test(Canopy_Density_Adj_Trail ~ Site)))
```

* Pair-wise comparisons for clarity
```{r}
canopy_dens_aov <- aov(data = canopy_dens, Canopy_Density ~ Site * Location * EcoCondition )
TukeyHSD(canopy_dens_aov)
```

# Diameter at Breast Height Analysis

* Read in Tree data

```{r}
papp_trees <- read_xlsx(path = "PappFieldData.xlsx", sheet = 1)
# Filter out the dead trees
papp_trees <- papp_trees %>% 
  filter(!grepl(pattern = "dead", papp_trees$Species))
# Add invasive/native status
papp_trees <- left_join(papp_trees, papp_species, by = "Species")

# Change order of ecological conditions
papp_trees$EcoCondition <- factor(papp_trees$EcoCondition, 
                                   levels = c("Disturbed", "Undisturbed", "Ecologically Significant"))
```

## Visualizing the Data

* Plot DBH using facets to split by site and color to separate by ecocondition

```{r}
ggplot(data = papp_trees, aes(x = Location, y = DBH, fill = EcoCondition)) +
  geom_boxplot() +
  facet_grid(Site~.) +
  ylab("Diameter at Breast Height (DBH)") + 
  scale_fill_discrete(name = "Ecological\nCondition",
                      breaks = c("Disturbed", "Undisturbed", "Ecologically Significant"),
                      labels = c("Disturbed", "Undisturbed", "Ecologically\nSignificant")) +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-4-DBH-Loc-Site.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

## Analysis of Variance

* Perform an ANOVA to check for differences in DBH at Sites, Locations, and EcoConditions

```{r}
DBH_lm_all <- lm(data = papp_trees, DBH ~ Site * Location * EcoCondition )
kable(anova(DBH_lm_all), digits = 3, format = "pandoc", caption = "ANOVA table of DBH")
```

* Pair-wise comparisons for clarity
```{r}
DBH_aov <- aov(data = papp_trees, DBH ~ Site * Location * EcoCondition )
TukeyHSD(DBH_aov)
```

# Analysis of Total Plant Species Richness based on Ground Cover data

In plots with a radius of 5 m, all plants were identified to species and an estiamte of the percent cover of the plot for each species was made. 
Each row of the data.frame represents a different species within a plot. 
Because of this, the number of rows for each plot, excluding the row recording bare soil, represnts the species richness of that plot.

* Load vegan

```{r}
library(vegan)
```

* Remove bare ground from ground cover data

```{r}
ground_veg <- papp_ground[ !(papp_ground$Species %in% c("Bare ground")), ]
```

* Look at number of plant species in each plot
```{r}
plants_tally <- 
  ground_veg %>%
  group_by(Site, Plot_Num, Location, EcoCondition) %>%
  tally()
```

* Look at number of native and invasive plants at each plot

```{r}
nat_inv_plants_tally <- 
  ground_veg %>%
  group_by(Site, Plot_Num, Location, EcoCondition, Inv_Nat) %>%
  tally()

# change flags for native and invasive
nat_inv_plants_tally$Inv_Nat <- 
  ifelse(nat_inv_plants_tally$Inv_Nat == "n", yes = "Native", no = "Invasive")
nat_inv_plants_tally$Inv_Nat <- factor(nat_inv_plants_tally$Inv_Nat,
                                       levels = c("Native", "Invasive"))

#Separate the Native Data
native_plants_tally <- nat_inv_plants_tally %>%
  filter(Inv_Nat == "Native")
```

## Visualizing the data

* Plot Species Richness using facets to split by site and color to separate by ecocondition

```{r}
ggplot(data = nat_inv_plants_tally, aes(x = Location, y = n, fill = EcoCondition)) +
  geom_boxplot() +
  facet_grid(Site~Inv_Nat) +
  ylab("Plant Species Richness") + 
  scale_fill_discrete(name = "Ecological\nCondition",
                      breaks = c("Disturbed", "Undisturbed", "Ecologically Significant"),
                      labels = c("Disturbed", "Undisturbed", "Ecologically\nSignificant")) +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-6-PlantRichness-Loc-Site.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```


## Statistical Analysis

* Perform an ANOVA to check for differences in Species Richness at Sites, Locations, and EcoConditions

```{r}
nat_inv_plant_tally_lm_all <- lm(data = nat_inv_plants_tally, n ~ Site * Location * EcoCondition * Inv_Nat )
kable(anova(nat_inv_plant_tally_lm_all), digits = 3, format = "pandoc", caption = "ANOVA table of Plant Species Richness")
```

* Pair-wise comparisons for clarity
```{r}
Species_Richness_aov <- aov(data = nat_inv_plants_tally, n ~ Site * Location * EcoCondition )
TukeyHSD(Species_Richness_aov)
```

# Analysis of Total Relative Percent Cover by Site, Location, EcoCondition, and Invasive/Native

* Look at plant data based on Percent Cover

```{r}
ground_veg2 <- ground_veg[, !(colnames(ground_veg) %in% c("Braun_Blaunquet"))]

# Calculate relative percent cover first
ground_veg2 <-
  ground_veg2 %>% 
  group_by(Plot_Num) %>%
  mutate(Plot_Tot_Cover = sum(Percent_Cover))
ground_veg2$Relative_Percent_Cover <-
  ground_veg2$Percent_Cover / ground_veg2$Plot_Tot_Cover

# Group the Data 
papp_ground_cover <- ground_veg2 %>%
  group_by(Site, Location, EcoCondition, Plot_Num)

# change flags for native and invasive
papp_ground_cover$Inv_Nat <- 
  ifelse(papp_ground_cover$Inv_Nat == "n", yes = "Native", no = "Invasive")
papp_ground_cover$Inv_Nat <- factor(papp_ground_cover$Inv_Nat,
                                       levels = c("Native", "Invasive"))

# Summarize the data by plots
papp_ground_cover_summary <- 
  papp_ground_cover %>%
  group_by(Site, Location, EcoCondition, Plot_Num, Inv_Nat) %>%
  summarise(Tot_Perc_Cover = sum(Percent_Cover),
            Tot_Rel_Perc_Cover = sum(Relative_Percent_Cover))

```

## Visualizing the data

* Plot Total Relative Percent Cover using facets to split by site and color to separate by ecocondition

```{r}
ggplot(data = papp_ground_cover_summary, aes(x = Location, y = Tot_Rel_Perc_Cover, fill = EcoCondition)) +
  geom_boxplot() +
  facet_grid(Site~Inv_Nat) +
  ylab("Total Relative Percent Cover") + 
  scale_fill_discrete(name = "Ecological\nCondition",
                      breaks = c("Disturbed", "Undisturbed", "Ecologically Significant"),
                      labels = c("Disturbed", "Undisturbed", "Ecologically\nSignificant")) +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-7-RelCov-Loc-Site.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

## Statistical Analysis

* Perform an ANOVA to check for differences in Total Relative Percent Cover at Sites, Locations, and EcoConditions

```{r}
inv_nat_tot_rel_cover_lm_all <- lm(data = papp_ground_cover_summary, Tot_Rel_Perc_Cover ~ Site * Location * EcoCondition * Inv_Nat )
kable(anova(inv_nat_tot_rel_cover_lm_all), digits = 3, format = "pandoc", caption = "ANOVA table of Total Relative Percent Cover")
```

* Pair-wise comparisons for clarity
```{r}
Tot_Rel_Cov_aov <- aov(data = papp_ground_cover_summary, Tot_Rel_Perc_Cover ~ Site * Location * EcoCondition )
TukeyHSD(Tot_Rel_Cov_aov)
```

# Analysis of Total Percent Cover by Site, Location, EcoCondition, and Invasive/Native

## Visualizing the data

* Plot Total Percent Cover using facets to split by site and color to separate by ecocondition

```{r}
ggplot(data = papp_ground_cover_summary, aes(x = Location, y = Tot_Perc_Cover, fill = EcoCondition)) +
  geom_boxplot() +
  facet_grid(Site~Inv_Nat) +
  ylab("Total Percent Cover") + 
  scale_fill_discrete(name = "Ecological\nCondition",
                      breaks = c("Disturbed", "Undisturbed", "Ecologically Significant"),
                      labels = c("Disturbed", "Undisturbed", "Ecologically\nSignificant")) +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-8-Cov-Loc-Site.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

## Statistical Analysis

* Perform an ANOVA to check for differences in Total Percent Cover at Sites, Locations, and EcoConditions

```{r}
inv_nat_tot_cover_lm_all <- lm(data = papp_ground_cover_summary, Tot_Perc_Cover ~ Site * Location * EcoCondition * Inv_Nat )
kable(anova(inv_nat_tot_cover_lm_all), digits = 3, format = "pandoc", caption = "ANOVA table of Total Percent Cover")
```

*Pair-wise comparisons for clarity
```{r}
Tot_Perc_Cover_aov <- aov(data = papp_ground_cover_summary, Tot_Perc_Cover ~ Site * Location * EcoCondition )
TukeyHSD(Tot_Perc_Cover_aov)
```

# Shannon Index of Dissimilarity for Tree Data

```{r}
tree_species_tally <- 
  papp_trees %>%
  group_by(Species, Site, Plot_Num, Location, EcoCondition) %>%
  tally()
```

* Use 'Spread' function to match the model set by BCI data

```{r}
tree_spread <- tree_species_tally %>%
  spread(Species, n, convert = TRUE)
```

* Remove Unnecessary Variables, Set 'N/A' values to Zero, Set Row Names to Plot Number

```{r}
trees_BCI <- tree_spread[, !(colnames(tree_spread) %in% c("Site", "Location", "EcoCondition"))] ;
  trees_BCI[is.na(trees_BCI <- trees_BCI)] <- 0 ;
   
trees_BCI <- data.frame(trees_BCI, row.names = 1)
```

* Use 'diversity' function to find the Shannon Values for each plot and calculate the mean

```{r}
shannon_trees <- diversity(trees_BCI)
canopy_dens$shannon_trees <- shannon_trees
```

## Visualize the Data

* Plot values using a boxplot for Shannon Index

```{r}
ggplot(data = canopy_dens, aes(x = Location, y = shannon_trees, fill = EcoCondition)) +
  geom_boxplot() +
  facet_grid(Site ~ .)
```

## Analysis of Variance

* Perform an ANOVA to check for differences in Shannon Index of Trees at Sites, Locations, and EcoConditions

```{r}
canopy_dens_sh_all <- lm(data = canopy_dens, shannon_trees ~ Site * Location * EcoCondition )
kable(anova(canopy_dens_sh_all), digits = 3, format = "pandoc", caption = "ANOVA table of Shannon Index of Trees")
```

* Pair-wise comparisons for clarity
```{r}
Shan_trees_aov <- aov(data = canopy_dens, shannon_trees ~ Site * Location * EcoCondition )
TukeyHSD(Shan_trees_aov)
```
# Shannon Index of Dissimilarity for Plant Data

* Use 'Spread' function to match the model set by BCI data

```{r}
ground_spread <- ground_veg2 %>%
  select(Species, Relative_Percent_Cover, Plot_Num) %>%
  spread(Species, Relative_Percent_Cover, convert = TRUE)
```

* Remove Unnecessary Variables, Set 'N/A' values to Zero, Set Row Names to Plot Number Vaues

```{r}
ground_BCI <- ground_spread
ground_BCI[is.na(ground_BCI <- ground_BCI)] <- 0 
ground_BCI <- data.frame(ground_BCI, row.names = 1)
```

* Use 'diversity' function to find the Shannon Values for each plot and calculate the mean

```{r}
shannon_ground <- diversity(ground_BCI)
plants_tally$shannon_ground <- shannon_ground
```

## Visualizing the Data

* Plot Shannon Index using facets to split by site and color to separate by ecocondition

```{r}
ggplot(data = plants_tally, aes(x = Location, y = shannon_ground, fill = EcoCondition)) +
  geom_boxplot()   +
  facet_grid(Site~.) +
  ylab("Shannon Index") + 
  scale_fill_discrete(name = "Ecological\nCondition",
                      breaks = c("Disturbed", "Undisturbed", "Ecologically Significant"),
                      labels = c("Disturbed", "Undisturbed", "Ecologically\nSignificant")) +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-9-Shan-Loc-Site.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

## Statistical Analysis

* Perform an ANOVA to check for differences in Shannon Index of Plants at Sites, Locations, and EcoConditions

```{r}
papp_ground_sh_all <- lm(data = plants_tally, shannon_ground ~ Site * Location * EcoCondition )
kable(anova(papp_ground_sh_all), digits = 3, format = "pandoc", caption = "ANOVA table of Shannon Index of Plants")
```

* Pair-wise comparisons for clarity
```{r}
Shan_ground_aov <- aov(data = plants_tally, shannon_ground ~ Site * Location * EcoCondition )
TukeyHSD(Shan_ground_aov)
```

# Shannon Index of Dissimilarity for Native Plant Data

* Use 'Spread' function to match the model set by BCI data

```{r}
native_ground_spread <- ground_veg2 %>%
  filter(Inv_Nat == "n") %>%
  select(Species, Relative_Percent_Cover, Plot_Num) %>%
  spread(Species, Relative_Percent_Cover, convert = TRUE)
```

* Remove Unnecessary Variables, Set 'N/A' values to Zero, Set Row Names to Plot Number Vaues

```{r}
native_ground_BCI <- native_ground_spread
native_ground_BCI[is.na(native_ground_BCI <- native_ground_BCI)] <- 0 
native_ground_BCI <- data.frame(native_ground_BCI, row.names = 1)
```

* Use 'diversity' function to find the Shannon Values for each plot and calculate the mean

```{r}
native_shannon_ground <- diversity(native_ground_BCI)
plants_tally$native_shannon_ground <- native_shannon_ground
```

## Visualizing the Data

* Plot Shannon Index using facets to split by site and color to separate by ecocondition

```{r}
ggplot(data = plants_tally, aes(x = Location, y = native_shannon_ground, fill = EcoCondition)) +
  geom_boxplot()   +
  facet_grid(Site~.) +
  ylab("Shannon Index") + 
  scale_fill_discrete(name = "Ecological\nCondition",
                      breaks = c("Disturbed", "Undisturbed", "Ecologically Significant"),
                      labels = c("Disturbed", "Undisturbed", "Ecologically\nSignificant")) +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-10-Nat-Shan-Loc-Site.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

## Statistical Analysis

* Perform an ANOVA to check for differences in Shannon Index of Native Plants at Sites, Locations, and EcoConditions

```{r}
papp_ground_sh_all <- lm(data = plants_tally, native_shannon_ground ~ Site * Location * EcoCondition )
kable(anova(papp_ground_sh_all), digits = 3, format = "pandoc", caption = "ANOVA table of Shannon Index of Plants")
```

* Pair-wise comparisons for clarity
```{r}
Shan_nat_ground_aov <- aov(data = plants_tally, native_shannon_ground ~ Site * Location * EcoCondition )
TukeyHSD(Shan_nat_ground_aov)
```

# Bray-Curtis Index Tree Data

* Use `vegdist` function to find Bray-Curtis dissimilarity values for tree data

```{r}
beta_bray <- vegdist(trees_BCI, method = "bray")

# Reshape the beta diversity matrix from wide to long format
beta_bray_mat <- as.matrix(beta_bray)
beta_bray_mat[!lower.tri(beta_bray_mat)] <- NA
beta_bray_df <- as.data.frame(beta_bray_mat)
beta_bray_df$plot1 <- row.names(beta_bray_df)
beta_bray_df_long <- gather(beta_bray_df, "plot2", "beta_bray", 1:60)
beta_bray_df_long <- filter(beta_bray_df_long, !is.na(beta_bray))

# Merge beta_bray_df_long with data.frame columns for plot characteristics
beta_bray_df_long <- inner_join(beta_bray_df_long, select(canopy_dens, Site:EcoCondition), by = c("plot1" = "Plot_Num"))
beta_bray_df_long <- inner_join(beta_bray_df_long, select(canopy_dens, Site:EcoCondition), by = c("plot2" = "Plot_Num"))

```

## Visualizing the Data

```{r}
ggplot(data = filter(beta_bray_df_long, Site.x == Site.y), aes(x = Site.x, y = beta_bray)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-11-Bray-Tree-Site.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

Use a simple t-test to see if there are significant differences in turnover among plots between sites

```{r}
beta_bray_temp <- filter(beta_bray_df_long, Site.x == Site.y)
t.test(data = beta_bray_temp, beta_bray~Site.x)
```

Now let's look for differences also based on edge vs interior

```{r}
beta_bray_temp <- filter(beta_bray_df_long, Site.x == Site.y, Location.x == Location.y)
ggplot(data = beta_bray_temp, aes(x = Site.x, y = beta_bray, fill = Location.x)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-12-Bray-Tree-Loc.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

Perform an ANOVA on these data to see if there's a difference between sites and edge/interior

```{r}
kable(anova(lm(data = beta_bray_temp, beta_bray ~ Site.x * Location.x)), digits = 3, format = "pandoc", caption = "ANOVA table")
```
Nope

* Pair-wise comparisons for clarity
```{r}
Bray_Tree_aov <- aov(data = beta_bray_temp, beta_bray ~ Site.x * Location.x )
TukeyHSD(Bray_Tree_aov)
```


Now let's check turn-over between edges and interiors

```{r}
beta_bray_temp <- filter(beta_bray_df_long, Site.x == Site.y, Location.x != Location.y)
ggplot(data = beta_bray_temp, aes(x = Site.x, y = beta_bray)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-13-Bray-Tree-Turn-Loc.tiff", device = "tiff", width = 6.5, height = 4, units = "in")

t.test(data = beta_bray_temp, beta_bray~Site.x)

```

# Bray-Curtis Index Plant Data

## Bray-Curtis

* Use `vegdist` function to find Bray-Curtis dissimilarity values for ground vegetation

```{r}
ground_beta_bray <- vegdist(ground_BCI, method = "bray")

# Reshape the beta diversity matrix from wide to long format
ground_beta_bray_mat <- as.matrix(ground_beta_bray)
ground_beta_bray_mat[!lower.tri(ground_beta_bray_mat)] <- NA
ground_beta_bray_df <- as.data.frame(ground_beta_bray_mat)
ground_beta_bray_df$plot1 <- row.names(ground_beta_bray_df)
ground_beta_bray_df_long <- gather(ground_beta_bray_df, "plot2", "ground_beta_bray", 1:60)
ground_beta_bray_df_long <- filter(ground_beta_bray_df_long, !is.na(ground_beta_bray))

# Merge ground_beta_bray_df_long with data.frame columns for plot characteristics
ground_beta_bray_df_long <- inner_join(ground_beta_bray_df_long, select(canopy_dens, Site:EcoCondition), by = c("plot1" = "Plot_Num"))
ground_beta_bray_df_long <- inner_join(ground_beta_bray_df_long, select(canopy_dens, Site:EcoCondition), by = c("plot2" = "Plot_Num"))

```

```{r}
ggplot(data = filter(ground_beta_bray_df_long, Site.x == Site.y), aes(x = Site.x, y = ground_beta_bray)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-15-Bray-Veg.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

Use a simple t-test to see if there are significant differences in turnover among plots between sites

```{r}
ground_beta_bray_temp <- filter(ground_beta_bray_df_long, Site.x == Site.y)
t.test(data = ground_beta_bray_temp, ground_beta_bray~Site.x)
```

Now let's look for differences also based on edge vs interior

```{r}
ground_beta_bray_temp <- filter(ground_beta_bray_df_long, Site.x == Site.y, Location.x == Location.y)
ggplot(data = ground_beta_bray_temp, aes(x = Site.x, y = ground_beta_bray, fill = Location.x)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-16-Bray-Veg-Loc.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

Perform an ANOVA on these data to see if there's a difference between sites and edge/interior

```{r}
kable(anova(lm(data = ground_beta_bray_temp, ground_beta_bray ~ Site.x * Location.x)), digits = 3, format = "pandoc", caption = "ANOVA table of Bray-Curtis of Total Ground Vegetation")
```

* Pair-wise comparisons for clarity
```{r}
Bray_Ground_aov <- aov(data = ground_beta_bray_temp, ground_beta_bray~ Site.x * Location.x )
TukeyHSD(Bray_Ground_aov)
```

Now let's check turn-over between edges and interiors

```{r}
ground_beta_bray_temp <- filter(ground_beta_bray_df_long, Site.x == Site.y, Location.x != Location.y)
ggplot(data = ground_beta_bray_temp, aes(x = Site.x, y = ground_beta_bray)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-17-Bray-Veg-Turn-Loc.tiff", device = "tiff", width = 6.5, height = 4, units = "in")

t.test(data = ground_beta_bray_temp, ground_beta_bray~Site.x)

```

# Bray-Curtis Index Native Plant Data

## Bray-Curtis

* Use `vegdist` function to find Bray-Curtis dissimilarity values for native ground vegetation

```{r}
native_ground_beta_bray <- vegdist(native_ground_BCI, method = "bray")

# Reshape the beta diversity matrix from wide to long format
native_ground_beta_bray_mat <- as.matrix(native_ground_beta_bray)
native_ground_beta_bray_mat[!lower.tri(native_ground_beta_bray_mat)] <- NA
native_ground_beta_bray_df <- as.data.frame(native_ground_beta_bray_mat)
native_ground_beta_bray_df$plot1 <- row.names(native_ground_beta_bray_df)
native_ground_beta_bray_df_long <- gather(native_ground_beta_bray_df, "plot2", "native_ground_beta_bray", 1:60)
native_ground_beta_bray_df_long <- filter(native_ground_beta_bray_df_long, !is.na(native_ground_beta_bray))

# Merge native_ground_beta_bray_df_long with data.frame columns for plot characteristics
native_ground_beta_bray_df_long <- inner_join(native_ground_beta_bray_df_long, select(canopy_dens, Site:EcoCondition), by = c("plot1" = "Plot_Num"))
native_ground_beta_bray_df_long <- inner_join(native_ground_beta_bray_df_long, select(canopy_dens, Site:EcoCondition), by = c("plot2" = "Plot_Num"))

```

```{r}
ggplot(data = filter(native_ground_beta_bray_df_long, Site.x == Site.y), aes(x = Site.x, y = native_ground_beta_bray)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-19-Bray-Nat-Veg.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

Use a simple t-test to see if there are significant differences in turnover among plots between sites

```{r}
native_ground_beta_bray_temp <- filter(native_ground_beta_bray_df_long, Site.x == Site.y)
t.test(data = native_ground_beta_bray_temp, native_ground_beta_bray~Site.x)
```

Now let's look for differences also based on edge vs interior

```{r}
native_ground_beta_bray_temp <- filter(native_ground_beta_bray_df_long, Site.x == Site.y, Location.x == Location.y)
ggplot(data = native_ground_beta_bray_temp, aes(x = Site.x, y = native_ground_beta_bray, fill = Location.x)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-20-Bray-Nat-Veg-Loc.tiff", device = "tiff", width = 6.5, height = 4, units = "in")
```

Perform an ANOVA on these data to see if there's a difference between sites and edge/interior

```{r}
kable(anova(lm(data = native_ground_beta_bray_temp, native_ground_beta_bray ~ Site.x * Location.x)), digits = 3, format = "pandoc", caption = "ANOVA table of Bray-Curtis for Native Ground Vegetation")
```

* Pair-wise comparisons for clarity
```{r}
Bray_Nat_Ground_aov <- aov(data = native_ground_beta_bray_temp, native_ground_beta_bray~ Site.x * Location.x )
TukeyHSD(Bray_Nat_Ground_aov)
```

Now let's check turn-over between edges and interiors

```{r}
native_ground_beta_bray_temp <- filter(native_ground_beta_bray_df_long, Site.x == Site.y, Location.x != Location.y)
ggplot(data = native_ground_beta_bray_temp, aes(x = Site.x, y = native_ground_beta_bray)) +
  geom_boxplot() +
  ylab("Bray-Curtis Value") + 
  xlab("Site") +
  theme_bw(base_family = "Times")
ggsave("../manuscript/figures/Figure-21-Bray-Nat-Veg-Turn-Loc.tiff", device = "tiff", width = 6.5, height = 4, units = "in")


t.test(data = native_ground_beta_bray_temp, native_ground_beta_bray~Site.x)
```
